---
title: "Week 8b Markdown Assignment"
author: "Christopher"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
getwd()
setwd('C:/CODING/Math17/Week08/')
library('readxl')
Mysales2021 <- read_excel("Sales2021.xlsx")
Expected.Distribution <- Mysales2021$Percent/sum(Mysales2021$Percent)
Q1 <- chisq.test(Mysales2021$Q1, p = Expected.Distribution)
Q2 <- chisq.test(Mysales2021$Q2, p = Expected.Distribution)
Q3 <- chisq.test(Mysales2021$Q3, p = Expected.Distribution)
Sales2021a <- sprintf('Q1= xsq: %s p-value:%s Q2= xsq: %s p-value:%s Q3= xsq: %s p-value:%s',Q1$statistic,Q1$p.value,Q2$statistic,Q2$p.value,Q3$statistic,Q3$p.value)
```

Sales figures for the first three quarters of 2021 have been posted and what follows is a 
chi-squared analysis of the quarterly sales vs the expected Percent of sales.

```{r analysis results}
Sales2021a
```

H0: The sales for a given quarter was independent of the Percent expectations for the quarter.
Ha: The sales for a given quarter was dependent of the Percent expectations for the quarter.

Q1: p-value = 1 > 0.05 (Failed to reject the null hypothesis)
Q2: p-value = 1 > 0.05 (Failed to reject the null hypothesis)
Q3: p-value = 2.2e-16 < 0.05 (Reject the null hypothesis)

Conclusion: At 5% level of significance, from the data, there is sufficient evidence to 
conclude that Q1 & Q2 failed to meet the Percent of expected sales, however Q3 met the 
Percent of of expected sales for 2021.
